---
title: "Carl Zimmer Chromosome 3 SNP Burden Analysis"
author: "Jeff Mandell, Guannan Gong, Neng Wan, and Lukas Fuentes"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read in annotated SNP data.
This analysis will be run with two sets of SNP data. The first excludes intergenic and intronic SNPs, but keeps splice-site, but keeps UTR SNPs and SNPS within a neighborhood upstream/downstream of the exons, while the second set only includes SNPs annotated as exon/splice-site.
```{r, tidy = T}
  data = read.table("chr3_annotated_all.txt", sep = "\t", header=T, na.strings = ".", stringsAsFactors = F, quote='')
  #data = read.table("chr3_annotated_exon-splice.txt", sep = "\t", header=T, na.strings = ".", stringsAsFactors = F, quote = '')
```


## Print gene lists
Next, print out the ten genes with highest mutational burden. Also print out lists of genes with the most non-synonymous mutations and with the most uncommon non-synonymous mutations (< 5% max popualtion frequency).

```{r, tidy = T}
  count_table = table(data$hgnc_symbol)
  print(sort(count_table, decreasing = T)[1:10])

  # Repeat, with synonymous variants excluded
  nonsynonymous = data[is.na(data$ExonicFunc.ensGene) | data$ExonicFunc.ensGene != "synonymous SNV", ]
  nonsyn_table = table(nonsynonymous$hgnc_symbol)
  print(sort(nonsyn_table, decreasing = T)[1:10])
  

  # Repeat using uncommon variants only (max population frequency < 5% or unknown)
  nonsynonymous_uncommon = nonsynonymous[is.na(nonsynonymous$PopFreqMax) | nonsynonymous$PopFreqMax < 0.05,]
  nonsyn_uncommon_table = table(nonsynonymous_uncommon$hgnc_symbol)
  print(sort(nonsyn_uncommon_table, decreasing = T)[1:10])
```


## Re-run with gene length normalization
Print the lists of genes with the highest variant burden (all variants, non-synonymous, and uncommon non-synonymous).
```{r, cache = T}
  ens_hcgna = data$Gene.ensGene
  names(ens_hcgna) = data$hgnc_symbol
  
  lengths = data$gene_length
  names(lengths) = data$Gene.ensGene
  normed = count_table / lengths[ens_hcgna[names(count_table)]]
  print(sort(normed, decreasing = T)[1:10])
  
  nonsynonymous_normed = nonsyn_table / lengths[ens_hcgna[names(nonsyn_table)]]
  print(sort(nonsynonymous_normed, decreasing = T)[1:10])
  
  nonsyn_uncommon_normed = nonsyn_uncommon_table / lengths[ens_hcgna[names(nonsyn_uncommon_table)]]
  print(sort(nonsyn_uncommon_normed, decreasing = T)[1:10])
```

